The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-04-26 paper.qmd
Moved over more hand drawn images

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
#| label: fig-vistax
#| fig-cap: "Planned visualisation: Add an R coded visualisation that shows the 5 considerations for showing uncertainty in a plot according to @Kinkeldey2014. One that has the consideration, and one that doesn't (try and keep as much the same between visuals as possible)"
# Library
library(tidyverse)

# Generate Data
data_short <- tibble(x = rnorm(100, sd = 1),
            y = rnorm(100, sd = 2))
data_long <- data_short |>
  pivot_longer(everything())

# implicit + visually integral
ggplot(data_short, aes(x=x, y="X")) + geom_jitter() 
# explicit
ggplot(data_short, aes(x=x)) + geom_density()

# intrinsic + separable
# avg data
avg_data <- data_long |> 
  group_by(name) |>
  summarise(mean = mean(value),
            sd = sd(value))
# plot
ggplot(avg_data, aes(x=mean, y=name, colour=sd)) +
  geom_point()

# extrinsic
ggplot(data_long, aes(x=value, y=name)) +
  geom_violin() + 
  geom_point(avg_data, aes(x=mean))

# coincidence
## Basically all above plots

# adjacent
avg_data |>
  pivot_longer()
ggplot(avg_data, aes(x=mean, y=name, colour=sd)) +
  geom_point()

# static

# dynamic

![Planned visualisation: Add an R coded visualisation that shows the 5 considerations for showing uncertainty in a plot according to @Kinkeldey2014. One that has the consideration, and one that doesn't (try and keep as much the same between visuals as possible)](notice.jpeg){#fig-location width=50%}

@@

![Visualisation showing the difference ](extractvsdecision.jpeg){#fig-extract}

@@

>>>>

![Visualisation of @Kinkeldey2014 uncertainty visualisation taxonomy. It can be seen that most of the categories can be considered a change to the typical error bar chart. Many of the distinctions conflate changes of the input data and statistics calculated and with visual changes, highlighting how defining uncertainty and visualisation together conflates several elements of a graphic.](uncertvistax.jpeg){#fig-location width=70%}

@@

![Visualisation of the process of transforming raw data to a visualisation insight. Summarising the data using common statistics removes information from our data. Once the data is in a format that represents the varibales of interest, organising that information in a visualisation makes it easier to extract insights from these statistics and convert them to insights. If information is removed at the statistics stage, it cannot be added back in at the visualisation stage, as it is meerly an efficient technique to organise and present information to enable a large range of insights. ](visprocess.jpeg){#fig-location width=70%}

@@

```{r}
